\thetitle

### 2020-11-08 sun_phases/spin_qudit_tomo.tex
updated some discussions

\renewcommand{\O}{\mathcal{O}}
\newcommand{\Q}{\mathcal{Q}}
\newcommand{\R}{\mathcal{R}}
\renewcommand{\S}{\mathcal{S}}

@@

A set of measurement axes $V=\set{\v v}$ nominally induces a set of projectors $\set{\Pi_{\v v\mu}}$ that will be measured in an experiment.
By a simple change of basis, measuring these projectors is equivalent to measuring the set of transition operators $\v T_V\equiv\set{T_{\v v\ell,0}}$.
By flattening the operators in $\v T_V$ into column vectors $\oket{T_{\v v\ell,0}}$, we can construct the {\it measurement matrix}
\begin{align}
  M_V \equiv \sum_{\v v,\ell} \oket{T_{\v v\ell,0}} \bra{\v v\ell},
  \label{eq:meas_mat}
\end{align}
A necessary and sufficient condition for $V$ to allow for full state tomography is that the measured transition operators in $\v T_V$, or equivalently the columns of $M_V$, span the entire ($d^2$-dimensional) space of operators on a $d$-level spin qudit.
In this case $M_V$ must be full rank, with $d^2$ nonzero singular values.
Indexing these singular values $M^V_k$ and the corresponding (normalized) right singular vectors $\v x^V_k\equiv(x^V_{k,1},x^V_{k,2},\cdots)$ by an integer $k\in\set{1,2,\cdots,d^2}$, we can construct the orthonormal qudit operators
\begin{align}
  Q^V_k \equiv \sum_j q^V_{kj} T_j,
  &&
  \v q^V_k \equiv \f{\v x^V_k}{M^V_k},
\end{align}
where for shorthand we use a combined index $j=\p{\v v_j,\ell_j}$ to specify both a measurement axis $\v v_j$ and a degree $\ell_j$, which identify the transition operator $T_j\equiv T_{\v v_j\ell_j,0}$.
These operators allow us to expand any state $\rho$ of a $d$-level spin qudit in the form
\begin{align}
  \rho = \sum_{k=1}^{d^2} \rho_k^V Q_k^V,
  &&
  \rho_k^V \equiv \bk{Q^V_k}_\rho.
\end{align}
Given empirical estimates $\tilde T_j$ of the expectation values $\bk{T_j}_\rho$, an empirical estimate $\tilde\rho_V$ of $\rho$ is then
\begin{align}
  \tilde\rho_V \equiv \sum_k \tilde\rho^V_k Q^V_k,
  \label{eq:reconstructed_state}
\end{align}
where
\begin{align}
  \tilde\rho^V_k \equiv \sum_j q^V_{kj} \tilde T_j
  \approx \sum_j q^V_{kj} \bk{T_j}_\rho
  = \bk{Q^V_k}_\rho
  = \rho^V_k.
\end{align}
The singular values of $M_V$ allow us to make concrete statements about the statistical error between the empirical estimate $\tilde\rho_V$ and the true state $\rho$.
Assume, for example, that the estimates $\tilde T_j$ are equal to $\bk{T_j}_\rho$ up to uncorrelated noise with variance at most $\epsilon^2$:
\begin{align}
  \tilde T_j = \bk{T_j}_\rho + \epsilon_j,
  &&
  \bbk{\epsilon_j \epsilon_{j'}} \le \epsilon^2 \delta_{jj'},
  \label{eq:error_assumption}
\end{align}
where $\set{\epsilon_j}$ are independent random variables and we use the double brackets $\bbk{\cdot}$ to denote statistical averaging.
In this case, the mean squared error with which $\tilde\rho^V_k$ approximates $\rho^V_k$ is
\begin{align}
  \Bbk{\abs{\tilde\rho^V_k-\rho^V_k}^2}
  &= \Bbk{\p{\tilde\rho^V_k-\rho^V_k}^*\p{\tilde\rho^V_k-\rho^V_k}} \\
  &= \sum_{j,j'} \p{q^V_{kj}}^* q^V_{kj'}\,
  \bbk{\epsilon_j \epsilon_{j'}} \\
  &\le \sum_j \abs{q^V_{kj}}^2 \epsilon^2
  = \p{\f{\epsilon}{M^V_k}}^2.
\end{align}
Using the fact that the operators $Q^V_k$ are orthonormal, we can therefore bound the mean squared (Euclidean) distance (i.e.~the distance induced by the trace inner product) between $\tilde\rho_V$ and $\rho$ as
\begin{align}
  \Bbk{\norm{\tilde\rho_V-\rho}^2}
  = \sum_k \Bbk{\abs{\tilde\rho^V_k - \rho^V_k}^2}
  \le \epsilon^2 \S_V^2,
  \label{eq:bound_eps}
\end{align}
where
\begin{align}
  \S_V^2 \equiv \norm{M_V^{-1}}^2 = \sum_k\f1{\p{M^V_k}^2},
  \label{eq:scale}
\end{align}
is the squared (Hilbert-Schmidt, or Frobenius) norm of the right inverse $M_V^{-1}$ of $M_V$.
Note that $\S_V^2$ diverges if $M_V$ is singular, in which case measuring spin projections along all axes in $V$ does not provide sufficient information to reconstruct arbitrary quantum states.

Computing $\S_V^2$ to certify the bound in Eq.~\eqref{eq:bound_eps} requires constructing the measurement matrix $M_V$ and finding its singular values.
This complexity of this task can be greatly reduced by the fact that the degree $\ell$ of the transition operator $T_{\ell,0}$ is preserved under rotations, which implies that the unitary
\begin{align}
  U \equiv \sum_{\ell,m} \ket{\ell m} \obra{T_{\ell m}},
\end{align}
block-diagonalizes the measurement matrix:
\begin{align}
  U M_V = \sum_\ell \op{\ell} \otimes M_{V\ell},
\end{align}
where the $\p{2\ell+1}\times\abs{V}$ blocks are
\begin{align}
  M_{V\ell} \equiv \sum_{m,\v v} \ket{m} \obk{T_{\ell m}|T_{\v v\ell,0}} \bra{\v v}
  = \sum_{m,\v v} D^\ell_{m,0}\p{\v v} \op{m}{\v v}.
\end{align}
As the singular values of $M_V$ are invariant under unitary transformations, it follows that
\begin{align}
  \S_V^2 = \sum_\ell \S_{V\ell}^2,
  &&
  \S_{V\ell}^2 \equiv \norm{M_{V\ell}^{-1}}^2,
\end{align}
where $M_{V\ell}^{-1}$ is the right inverse of $M_{V\ell}$.
Constructing the block $M_{V\ell}$ and computing its singular values takes at most $O(\abs{V}d^2)$ time (see Appendix \ref{sec:compute} for additional details on efficiently computing $\S_{V\ell}$ [\red{todo: write tutorial}]).
Altogether, if we assume that $\abs{V}\sim d$, then computing $\S_V^2$ to certify the bound in Eq.~\eqref{eq:bound_eps} takes $O(d^4)$ serial or $O(d^3)$ parallel runtime (see Figure [\red{REF}]).
[\red{todo: add figure showing runtimes as a function of $d$ with $\abs{V}=2d-1$; modify text appropriately after making the figure}].
For comparison, we note that existing algorithms for signal recovery on the sphere \cite{mcewen2011novel, rauhut2011sparse, alem2012sparse, khalid2014optimaldimensionality} similarly have $O(d^3)$ parallel runtime with $\abs{V}\gtrsim d$.

The assumption that observables can be estimated up to uncorrelated noise with maximal variance $\epsilon^2$, summarized by Eq.~\eqref{eq:error_assumption}, is reasonable when measurement error is dominated by experimental sources of noise.
However, this assumption breaks down when measurement error is limited by fundamental quantum (shot) noise.
We relax the assumption of Eq.~\eqref{eq:error_assumption} and derive more fundamental error bounds in Appendix \ref{sec:error}.
Therein, we find that if measurement error is dominated by shot noise, then the estimate $\tilde\rho_V$ built from $n$ independent measurements of spin projection along every axis $\v v\in V$ approximates the true state $\rho$ to within a mean squared distance
\begin{align}
  \Bbk{\norm{\tilde\rho_V-\rho}^2}
  \le \epsilon_V^2
  \equiv \f1{nd} \sum_{\ell>0} \p{2\ell+1} \gamma_\ell^2 \S_{V\ell},
  \label{eq:bound}
\end{align}
where
\begin{align}
  \gamma_\ell
  \equiv \f{\max_\mu\tau_{\ell\mu} - \min_\mu\tau_{\ell\mu}}{2}
\end{align}
is defined in terms of the Clebsch-Gordan coefficients
\begin{align}
  \tau_{\ell\mu} \equiv \bk{s\mu;\ell,0|s\mu}.
\end{align}
For reference, we note that $\epsilon_V^2<\S_V^2/n$.

@@

The ability to certify a statistical error bound on the empirical estimate $\tilde\rho_V$ of an unknown quantum state $\rho$ motivates the following simple protocol for spin qudit tomography: select a random set of measurement axes $V$ by uniformly sampling points on the sphere \footnote{To sample a point $\p{\alpha,\beta}$ from the uniform distribution on the sphere (with azimuthal angle $\alpha$ and polar angle $\beta$), you can sample a point $\p{a,b}\in[0,1]\times[0,1]$ from the uniform distribution on the unit square, and then set $\alpha=2\pi a$ and $\beta=\arccos\p{1-2b}$.}, and optimize (using any standard minimization algorithm) the $2\abs{V}$ parameters in $V$ (two angles for each point $\v v\in V$) to minimize the error scale $\S_V$ defined in Eq.~\eqref{eq:scale}.
If $\abs{V}$ is too large for optimization to be practical, one can instead generate many random sets of measurement axes, and then choose the set with the smallest error scale $\S_V$.
After choosing a set of measurement axes $V$, experimentally characterize expectation values of spin projections, $\bk{\Pi_{\v v\mu}}_\rho$, which determine the expectation values $\bk{T_{\v v\ell,0}}_\rho$.
If the experimental estimates of $\bk{T_{\v v\ell,0}}_\rho$ are accurate to within an error $\lesssim\epsilon$, then the state $\tilde\rho_V$ defined in Eq.~\eqref{eq:reconstructed_state} is an approximation of the true state $\rho$ to within a (Euclidean) distance $\lesssim\epsilon\S_V$.
Furthermore, if the empirical estimate $\tilde\rho_V$ has negative eigenvalues, its distance from $\rho$ can be further reduced with maximum-likelihood corrections \cite{smolin2012efficient}.

@@

As a last point, we mention that the error scale $\S_V$ can also evaluate the ``quality'' of measurement axes $V$ chosen with existing methods for sparse signal recovery on the sphere \cite{rauhut2011sparse, alem2012sparse}.
In practice, we find that uniform random sampling outperforms these methods, in the sense that it generates low-error-scale sets of measurement axes with higher probability.
Two observations may explain the poor performance of these sparse signal recovery methods.
First, spin qudit tomography essentially boils down to recovering sparse signals with a particular harmonic structure, namely a fixed degree $\ell$.
The existing sparse signal recovery algorithms do not assume or account for any harmonic structure, and may therefore perform worse than expected on this particular subset of signals.
Second, spin qudit tomography requires recovering signals that are band-limited at degree $d$, but have $O(d)$ non-zero harmonic coefficients (out of $d^2$ coefficients in total).
The existing sparse signal recovery methods, on the other hand, may only perform well for signals with sparsity much smaller than $d$.
In any case, the fact that uniform sampling outperforms these methods only simplifies our randomized tomography protocol.

@@

We conjecture, and numerically verify for $d\le200$, that measurements of spin projection along $2d-1$ axes suffice to reconstruct an arbitrary state of a $d$-level spin qudit (see Figure [\red{REF}]).
Even so, $2d-1$ measurement axes may not be the ``optimal'' number to use.
Increasing the number of measurement axes generally decreases the error scale $\S_V$, but at the cost of having to estimate more observables.
If we fix the total number of measurements (or experimental trials), then having to estimate more observables means that fewer measurements are devoted to estimating any particular observable, which in turn increases the error $\epsilon$ on those estimates.
This trade-off begs the question: how many measurement axes should be chosen to perform spin qudit tomography?

The reconstruction error bound in Eq.~\eqref{eq:bound_eps} provides a straightforward answer: the number of measurement axes should be chosen to minimize the product $\epsilon^2\S_V^2$, which bounds the expected error in the estimate $\tilde\rho_V$ of an unknown state $\rho$.
If there are a total of $N$ spin projection measurements onto $r$ quantization axes, then $n=N/r$ measurements are used to estimate spin projection onto any given axis.
The variance in estimates of spin projection is then $\epsilon^2\sim1/n\sim r$ at a fixed total number of measurements $N$.
Altogether, the number of measurement axes should be chosen in such a way as to minimize $r\S_V^2$ with $\abs{V}=r$.

@@

Rather than the statistical errors $\epsilon_{i\mu} \equiv \tilde\Pi_{i\mu} - \bk{\Pi_{i\mu}}_\rho$ in the estimates $\tilde\Pi_{i\mu}$ of the projectors $\Pi_{i\mu}$, we now consider the statistical errors $\epsilon_{i\ell} \equiv \tilde T_{i\ell} - \bk{T_{i\ell}}_\rho$ in the estimates $\tilde T_{i\ell}$ of the transition operators $T_{i\ell} \equiv T_{\v v_i\ell,0}$.
We can expand the transition operators $T_{i\ell}$ as a sum over projectors $\Pi_{i\mu}$ as
\begin{align}
  T_{i\ell} = \sum_\mu t_{\ell\mu} \Pi_{i\mu},
  &&
  t_{\ell\mu} \equiv \bk{\mu|T_{\ell,0}|\mu},
\end{align}
and likewise $\tilde T_{i\ell} = \sum_\mu t_{\ell\mu} \tilde\Pi_{i\mu}$.
The covariances between statistical errors in the transition operator basis are then
\begin{align}
  \bbk{\epsilon_{i\ell} \epsilon_{i'\ell'}}
  = \sum_{\mu,\mu'} t_{\ell\mu} t_{\ell'\mu'}
  \bbk{\epsilon_{i\mu} \epsilon_{i'\mu'}}
  = \delta_{ii'} \times \f1n \sum_{\mu,\mu'} t_{\ell\mu} t_{\ell'\mu'}
  \sp{\delta_{\mu\mu'} p_\mu^{(i)} - p_\mu^{(i)} p_{\mu'}^{(i)}}.
\end{align}
In order to establish an upper bound on reconstruction error for an unknown quantum state, we need an upper bound on the variance $\bbk{\epsilon_{i\ell}^2}$ in particular.
To this end, for any multinomial probability distribution $p\equiv\p{p_1,p_2,\cdots,p_d}$ we define the function
\begin{align}
  g_\ell\p{p} \equiv \sum_{\mu,\mu'} t_{\ell\mu} t_{\ell\mu'}
  \p{\delta_{\mu\mu'} p_\mu - p_\mu p_{\mu'}}
  = \sum_\mu t_{\ell\mu}^2 p_\mu - \p{\sum_\mu t_{\ell\mu} p_\mu}^2
  = \sigma_p^2\p{t_\ell},
\end{align}
where $t_\ell \equiv \p{t_{\ell,-s},t_{\ell,-s+1},\cdots,t_{\ell s}}$ is a vector of all $t_{\ell\mu}$, and
\begin{align}
  \sigma_p^2\p{X}
  \equiv \sum_\mu p_\mu \p{X_\mu - \sum_\nu p_\nu X_\nu}^2
  = \sum_\mu p_\mu X_\mu^2 - \p{\sum_\mu p_\mu X_\mu}^2
\end{align}
is the variance of $X$ when sampled according to the probability distribution $p$.
By expanding
\begin{align}
  t_{\ell\mu} = \sqrt{\f{2\ell+1}{d}}\, \tau_{\ell\mu},
  &&
  \tau_{\ell\mu} \equiv \bk{s\mu;\ell,0|s\mu},
\end{align}
we can simplify
\begin{align}
  g_\ell\p{p} = \f{2\ell+1}{d} \times \sigma_p^2\p{\tau_\ell}.
\end{align}
We now seek to maximize $g_\ell\p{p}$, which will provide an upper bound for $\bbk{\epsilon_{i\ell}^2}$.
In general, variance $\sigma_p^2\p{X}$ is maximal when $p$ has equal weight on the largest and smallest values of $X$, which implies that
\begin{align}
  \sigma_p^2\p{\tau_\ell} \le \gamma_\ell^2,
  &&
  \gamma_\ell \equiv \f{\max\tau_\ell - \min\tau_\ell}{2}.
\end{align}
Altogether, using additionally the fact that $g_0\p{p}=0$, we can bound
\begin{align}
  \bbk{\epsilon_{i\ell}^2}
  = \f1n \times g_\ell\p{p^{(i)}}
  \le \f{2\ell+1}{nd} \times \gamma_\ell^2
  \times \p{1 - \delta_{\ell,0}}.
  \label{eq:variance_bound}
\end{align}
Note that the bound in Eq.~\eqref{eq:variance_bound} is tight, as equality is achieved by the mixed state
\begin{align}
  \rho^\star
  \equiv \f12 \p{\Pi_{\v v_i,\mu_{\t{max}}} + \Pi_{\v v_i,\mu_{\t{min}}}},
\end{align}
where $\mu_{\t{max}}$ ($\mu_{\t{min}}$) is the index that maximizes (minimizes) $\tau_{\ell\mu}$.
In addition, the  Clebsch-Gordan coefficients $\tau_{\ell\mu}$ are at most $1$ in magnitude, which implies that $\gamma_\ell^2\le 1$.
However, the bound on $\gamma_\ell^2\le1$ is not tight in general, as we could have alternately bounded
\begin{align}
  g_\ell\p{p} \le \sum_\mu t_{\ell\mu}^2 p_\mu
  \le \sum_\mu t_{\ell\mu}^2 = 1,
\end{align}
which must also hold for the probability distribution $p^\star$ that maximizes $g_\ell$, i.e.
\begin{align}
  g_\ell\p{p^\star} = \f{2\ell+1}{d} \times \gamma_\ell^2 \le 1,
  &&
  \t{so}
  &&
  \gamma_\ell^2 \le \min\Set{1, \f{d}{2\ell+1}}.
\end{align}

>>>>

\renewcommand{\O}{\mathcal{O}}
\newcommand{\Q}{\mathcal{Q}}
\renewcommand{\S}{\mathcal{S}}

@@

A set of measurement axes $V=\set{\v v}$ nominally induces a set of projectors $\set{\Pi_{\v v\mu}}$ that will be measured in an experiment.
By a simple change of basis, measuring these projectors is equivalent to measuring the set of transition operators $\v T_V\equiv\set{T_{\v v\ell,0}}$.
By flattening the operators in $\v T_V$ into column vectors $\oket{T_{\v v\ell,0}}$, we can construct the {\it measurement matrix}
\begin{align}
  M_V \equiv \sum_{\v v,\ell} \oket{T_{\v v\ell,0}} \bra{\v v\ell},
  \label{eq:meas_mat}
\end{align}
A necessary and sufficient condition for $V$ to allow for full state tomography is that the measured transition operators in $\v T_V$, or equivalently the columns of $M_V$, span the entire ($d^2$-dimensional) space of operators on a $d$-level spin qudit.
In this case $M_V$ must be full rank, with $d^2$ nonzero singular values.
Indexing these singular values $M^V_k$ and the corresponding (normalized) right singular vectors $\v x^V_k\equiv(x^V_{k,1},x^V_{k,2},\cdots)$ by an integer $k\in\set{1,2,\cdots,d^2}$, we can construct the orthonormal qudit operators
\begin{align}
  Q^V_k \equiv \sum_j q^V_{kj} T_j,
  &&
  \v q^V_k \equiv \f{\v x^V_k}{M^V_k},
\end{align}
where for shorthand we use a combined index $j=\p{\v v_j,\ell_j}$ to specify both a measurement axis $\v v_j$ and a degree $\ell_j$, which identify the transition operator $T_j\equiv T_{\v v_j\ell_j,0}$.
These operators allow us to expand any state $\rho$ of a $d$-level spin qudit in the form
\begin{align}
  \rho = \sum_{k=1}^{d^2} \rho_k^V Q_k^V,
  &&
  \rho_k^V \equiv \bk{Q^V_k}_\rho.
\end{align}
Given empirical estimates $\tilde T_j$ of the expectation values $\bk{T_j}_\rho$, an empirical estimate $\tilde\rho_V$ of $\rho$ is then
\begin{align}
  \tilde\rho_V \equiv \sum_k \tilde\rho^V_k Q^V_k,
  \label{eq:reconstructed_state}
\end{align}
where
\begin{align}
  \tilde\rho^V_k \equiv \sum_j q^V_{kj} \tilde T_j
  \approx \sum_j q^V_{kj} \bk{T_j}_\rho
  = \bk{Q^V_k}_\rho
  = \rho^V_k.
\end{align}
The singular values of $M_V$ allow us to make concrete statements about the statistical error between the empirical estimate $\tilde\rho_V$ and the true state $\rho$.
Assume, for example, that the estimates $\tilde T_j$ are equal to $\bk{T_j}_\rho$ up to uncorrelated noise with variance at most $\epsilon^2$:
\begin{align}
  \tilde T_j = \bk{T_j}_\rho + \epsilon_j,
  &&
  \bbk{\epsilon_j \epsilon_{j'}} \le \epsilon^2 \delta_{jj'},
  \label{eq:error_assumption}
\end{align}
where $\set{\epsilon_j}$ are independent random variables and we use the double brackets $\bbk{\cdot}$ to denote statistical averaging.
In this case, the mean squared error with which $\tilde\rho^V_k$ approximates $\rho^V_k$ is
\begin{align}
  \Bbk{\abs{\tilde\rho^V_k-\rho^V_k}^2}
  &= \Bbk{\p{\tilde\rho^V_k-\rho^V_k}^*\p{\tilde\rho^V_k-\rho^V_k}} \\
  &= \sum_{j,j'} \p{q^V_{kj}}^* q^V_{kj'}\,
  \bbk{\epsilon_j \epsilon_{j'}} \\
  &\le \sum_j \abs{q^V_{kj}}^2 \epsilon^2
  = \p{\f{\epsilon}{M^V_k}}^2.
\end{align}
Using the fact that the operators $Q^V_k$ are orthonormal, we can therefore bound the mean squared (Euclidean) distance (i.e.~the distance induced by the trace inner product) between $\tilde\rho_V$ and $\rho$ as
\begin{align}
  \Bbk{\norm{\tilde\rho_V-\rho}^2}
  = \sum_k \Bbk{\abs{\tilde\rho^V_k - \rho^V_k}^2}
  \le \epsilon^2 \S_V^2,
  \label{eq:bound_eps}
\end{align}
where
\begin{align}
  \S_V^2 \equiv \norm{M_V^{-1}}^2 = \sum_k\f1{\p{M^V_k}^2},
\end{align}
is the squared (Hilbert-Schmidt, or Frobenius) norm of the right inverse $M_V^{-1}$ of $M_V$.
Note that $\S_V^2$ diverges if $M_V$ is singular, in which case measuring spin projections along all axes in $V$ does not provide sufficient information to reconstruct arbitrary quantum states.

Computing $\S_V^2$ to certify the bound in Eq.~\eqref{eq:bound_eps} requires constructing the measurement matrix $M_V$ and finding its singular values.
This complexity of this task can be greatly reduced by the fact that the degree $\ell$ of the transition operator $T_{\ell,0}$ is preserved under rotations, which implies that the unitary
\begin{align}
  U \equiv \sum_{\ell,m} \ket{\ell m} \obra{T_{\ell m}},
\end{align}
block-diagonalizes the measurement matrix:
\begin{align}
  U M_V = \sum_\ell \op{\ell} \otimes M_{V\ell},
\end{align}
where the $\p{2\ell+1}\times\abs{V}$ blocks are
\begin{align}
  M_{V\ell} \equiv \sum_{m,\v v} \ket{m} \obk{T_{\ell m}|T_{\v v\ell,0}} \bra{\v v}
  = \sum_{m,\v v} D^\ell_{m,0}\p{\v v} \op{m}{\v v}.
\end{align}
As the singular values of $M_V$ are invariant under unitary transformations, it follows that
\begin{align}
  \S_V^2 = \sum_\ell \S_{V\ell}^2,
  &&
  \S_{V\ell}^2 \equiv \norm{M_{V\ell}^{-1}}^2,
\end{align}
where $M_{V\ell}^{-1}$ is the right inverse of $M_{V\ell}$.
Constructing the block $M_{V\ell}$ and computing its singular values takes at most $O(\abs{V}d^2)$ time (see Appendix \ref{sec:compute} for additional details on efficiently computing $\S_{V\ell}^2$ [\red{todo: write tutorial}]).
Altogether, if we assume that $\abs{V}\sim d$, then computing $\S_V^2$ to certify the bound in Eq.~\eqref{eq:bound_eps} takes $O(d^4)$ serial or $O(d^3)$ parallel runtime (see Figure [\red{REF}]).
[\red{todo: add figure showing runtimes as a function of $d$ with $\abs{V}=2d-1$; modify text appropriately after making the figure}].
For comparison, we note that existing algorithms for signal recovery on the sphere \cite{mcewen2011novel, rauhut2011sparse, alem2012sparse, khalid2014optimaldimensionality} similarly have $O(d^3)$ parallel runtime with $\abs{V}\gtrsim d$.

The assumption that observables can be estimated up to uncorrelated noise with maximal variance $\epsilon^2$, summarized by Eq.~\eqref{eq:error_assumption}, is reasonable when measurement error is dominated by experimental sources of noise.
However, this assumption breaks down when measurement error is limited by fundamental quantum (shot) noise.
We relax the assumption of Eq.~\eqref{eq:error_assumption} and derive more fundamental error bounds in Appendix \ref{sec:error}.
Therein, we find that if measurement error is dominated by shot noise, then the estimate $\tilde\rho_V$ built from $n$ independent measurements of spin projection along every axis $\v v\in V$ approximates the true state $\rho$ to within a mean squared distance $\epsilon_V^2/n$, i.e.
\begin{align}
  \Bbk{\norm{\tilde\rho_V-\rho}^2}
  \le \f{\epsilon_V^2}{n},
  \label{eq:bound}
\end{align}
where the {\it error scale} $\epsilon_V^2$ is
\begin{align}
  \epsilon_V^2 \equiv
  \f1{nd} \sum_{\ell>0} \p{2\ell+1} \gamma_\ell^2 \S_{V\ell}^2,
  \label{eq:scale}
\end{align}
and $\gamma_\ell$ is defined using Clebsch-Gordan coefficients:
\begin{align}
  \gamma_\ell
  \equiv \f{\max_\mu\tau_{\ell\mu} - \min_\mu\tau_{\ell\mu}}{2},
  &&
  \tau_{\ell\mu} \equiv \bk{s\mu;\ell,0|s\mu}.
\end{align}
For reference, we note that the error scale $\epsilon_V^2<\S_V^2$.

@@

The ability to certify a statistical error bound on the empirical estimate $\tilde\rho_V$ of an unknown quantum state $\rho$ motivates the following simple protocol for spin qudit tomography: select a random set of measurement axes $V$ by uniformly sampling points on the sphere \footnote{To sample a point $\p{\alpha,\beta}$ from the uniform distribution on the sphere (with azimuthal angle $\alpha$ and polar angle $\beta$), you can sample a point $\p{a,b}\in[0,1]\times[0,1]$ from the uniform distribution on the unit square, and then set $\alpha=2\pi a$ and $\beta=\arccos\p{1-2b}$.}, and optimize (using any standard minimization algorithm) the $2\abs{V}$ parameters in $V$ (two angles for each point $\v v\in V$) to minimize the error scale $\epsilon_V^2$ defined in Eq.~\eqref{eq:scale}.
If $\abs{V}$ is too large for such optimization  to be practical, one can instead generate many random sets of measurement axes, and then choose the set with the smallest error scale $\epsilon_V^2$.
After choosing a set of measurement axes $V$, experimentally characterize the expectation values $\bk{\Pi_{\v v\mu}}_\rho$ by making measuring spin projection along every axis $\v v\in V$.
These spin projections determine the expectation values $\bk{T_{\v v\ell,0}}_\rho$, whose estimates can be used to construct the state $\tilde\rho$ defined in Eq.~\eqref{eq:reconstructed_state}.
If spin projection was measured $n$ times along every axis, then the reconstructed state $\tilde\rho$ approximates the true state $\rho$ to within a mean squared (Euclidean) distance $\epsilon_V^2/n$.
Furthermore, if $\tilde\rho_V$ has negative eigenvalues, its distance from $\rho$ can be further reduced with maximum-likelihood corrections \cite{smolin2012efficient}.

@@

As a last point, we mention that the error scale $\epsilon_V^2$ can also be used to evaluate the ``quality'' of measurement axes $V$ chosen with existing methods for sparse signal recovery on the sphere \cite{rauhut2011sparse, alem2012sparse}.
In practice, we find that uniform random sampling outperforms these methods, in the sense that it generates low-error-scale sets of measurement axes with higher probability.
Two observations may explain the poor performance of these sparse signal recovery methods.
First, spin qudit tomography essentially boils down to recovering sparse signals with a particular harmonic structure, namely a fixed degree $\ell$.
The existing sparse signal recovery algorithms do not assume or account for any harmonic structure, and may therefore perform worse than expected on this particular subset of signals.
Second, spin qudit tomography requires recovering signals that are band-limited at degree $d$, but have $O(d)$ non-zero harmonic coefficients (out of $d^2$ coefficients in total).
The existing sparse signal recovery methods, on the other hand, may only perform well for signals with sparsity much smaller than $d$.
In any case, the fact that uniform sampling outperforms these methods only simplifies our randomized tomography protocol.

@@

We conjecture, and numerically verify for $d\le200$, that measurements of spin projection along $2d-1$ axes suffice to reconstruct an arbitrary state of a $d$-level spin qudit (see Figure [\red{REF}]).
Even so, $2d-1$ measurement axes may not be the ``optimal'' number to use.
Increasing the number of measurement axes generally decreases the error scale $\epsilon_V^2$, but at the cost of having to estimate more observables.
At a fixed total number of measurements, $N=n\abs{V}$, increasing the number of measurement axes $\abs{V}$ reduces the number of measurements $n$ along each axis $\v v\in V$.
This trade-off begs the question: how many measurement axes should be chosen to perform spin qudit tomography?

The reconstruction error bound in Eq.~\eqref{eq:bound} nominally provides a straightforward answer: at a fixed total number of measurements $N=n\abs{V}$, the number of measurement axes should be chosen to minimize $\epsilon_V^2/n\propto\epsilon_V^2\abs{V}$.

@@

Rather than the statistical errors $\epsilon_{i\mu} \equiv \tilde\Pi_{i\mu} - \bk{\Pi_{i\mu}}_\rho$ in the estimates $\tilde\Pi_{i\mu}$ of the projectors $\Pi_{i\mu}$, we now consider the statistical errors $\epsilon_{i\ell} \equiv \tilde T_{i\ell} - \bk{T_{i\ell}}_\rho$ in the estimates $\tilde T_{i\ell}$ of the transition operators $T_{i\ell} \equiv T_{\v v_i\ell,0}$.
We can expand the transition operators $T_{i\ell}$ as a sum over projectors $\Pi_{i\mu}$ as
\begin{align}
  T_{i\ell} = \sum_\mu t_{\ell\mu} \Pi_{i\mu},
  &&
  t_{\ell\mu} \equiv \bk{\mu|T_{\ell,0}|\mu},
\end{align}
and likewise $\tilde T_{i\ell} = \sum_\mu t_{\ell\mu} \tilde\Pi_{i\mu}$.
The covariances between statistical errors in the transition operator basis are then
\begin{align}
  \bbk{\epsilon_{i\ell} \epsilon_{i'\ell'}}
  = \sum_{\mu,\mu'} t_{\ell\mu} t_{\ell'\mu'}
  \bbk{\epsilon_{i\mu} \epsilon_{i'\mu'}}
  = \delta_{ii'} \times \f1n \sum_{\mu,\mu'} t_{\ell\mu} t_{\ell'\mu'}
  \sp{\delta_{\mu\mu'} p_\mu^{(i)} - p_\mu^{(i)} p_{\mu'}^{(i)}}.
\end{align}
In order to establish an upper bound on reconstruction error for an unknown quantum state, we need an upper bound on the variance $\bbk{\epsilon_{i\ell}^2}$ in particular.
To this end, for any multinomial probability distribution $p\equiv\p{p_1,p_2,\cdots,p_d}$ we define the function
\begin{align}
  g_\ell\p{p} \equiv \sum_{\mu,\mu'} t_{\ell\mu} t_{\ell\mu'}
  \p{\delta_{\mu\mu'} p_\mu - p_\mu p_{\mu'}}
  = \sum_\mu t_{\ell\mu}^2 p_\mu - \p{\sum_\mu t_{\ell\mu} p_\mu}^2
  = \sigma_p^2\p{t_\ell},
\end{align}
where $t_\ell \equiv \p{t_{\ell,-s},t_{\ell,-s+1},\cdots,t_{\ell s}}$ is a vector of all $t_{\ell\mu}$, and
\begin{align}
  \sigma_p^2\p{X}
  \equiv \sum_\mu p_\mu \p{X_\mu - \sum_\nu p_\nu X_\nu}^2
  = \sum_\mu p_\mu X_\mu^2 - \p{\sum_\mu p_\mu X_\mu}^2
\end{align}
is the variance of $X$ when sampled according to the probability distribution $p$.
By expanding
\begin{align}
  t_{\ell\mu} = \sqrt{\f{2\ell+1}{d}}\, \tau_{\ell\mu},
  &&
  \tau_{\ell\mu} \equiv \bk{s\mu;\ell,0|s\mu},
\end{align}
we can simplify
\begin{align}
  g_\ell\p{p} = \f{2\ell+1}{d} \times \sigma_p^2\p{\tau_\ell}.
\end{align}
We now seek to maximize $g_\ell\p{p}$, which will provide an upper bound for $\bbk{\epsilon_{i\ell}^2}$.
In general, variance $\sigma_p^2\p{X}$ is maximal when $p$ has equal weight on the largest and smallest values of $X$, which implies that
\begin{align}
  \sigma_p^2\p{\tau_\ell} \le \gamma_\ell^2,
  &&
  \gamma_\ell \equiv \f{\max\tau_\ell - \min\tau_\ell}{2}.
\end{align}
Altogether, using additionally the fact that $g_0\p{p}=0$, we can bound
\begin{align}
  \bbk{\epsilon_{i\ell}^2}
  = \f1n \times g_\ell\p{p^{(i)}}
  \le \f{2\ell+1}{nd} \times \gamma_\ell^2
  \times \p{1 - \delta_{\ell,0}}.
  \label{eq:variance_bound}
\end{align}
Note that the bound in Eq.~\eqref{eq:variance_bound} is tight, as equality is achieved by the mixed state
\begin{align}
  \rho^\star
  \equiv \f12 \p{\Pi_{\v v_i,\mu_{\t{max}}} + \Pi_{\v v_i,\mu_{\t{min}}}},
\end{align}
where $\mu_{\t{max}}$ ($\mu_{\t{min}}$) is the index that maximizes (minimizes) $\tau_{\ell\mu}$.
In addition, the  Clebsch-Gordan coefficients $\tau_{\ell\mu}$ are at most $1$ in magnitude, which implies that $\gamma_\ell^2\le 1$.
However, the bound on $\gamma_\ell^2\le1$ is not tight in general, as we could have alternatively bounded
\begin{align}
  g_\ell\p{p} \le \sum_\mu t_{\ell\mu}^2 p_\mu
  \le \sum_\mu t_{\ell\mu}^2 = 1,
\end{align}
which must also hold for the probability distribution $p^\star$ that maximizes $g_\ell$, i.e.
\begin{align}
  g_\ell\p{p^\star} = \f{2\ell+1}{d} \times \gamma_\ell^2 \le 1,
  &&
  \t{so}
  &&
  \gamma_\ell^2 \le \min\Set{1, \f{d}{2\ell+1}}.
\end{align}
